---
title: "2_CSISA_KVK_Trials_causal_rf_targeting"
format: html
fig-dpi: 300
fig-width: 8.88
fig-align: center
fig-height: 5

self-contained: true
author: Maxwell Mkondiwa
editor: visual
toc: true
toc-location: left
number-sections: true
execute: 
  message: false
  warning: false
  echo: true
---

# Introduction

In this notebook, I use a causal machine learning estimator, i.e., multi-armed causal random forest with augmented inverse propensity score weights (Athey et al 2019), to estimate conditional average treatment effects (CATES) for agronomic practices. These CATEs are estimated for each individual farm thereby providing personalized estimates of the potential effectiveness of the practices. I then use a debiased robust estimator in a policy tree optimization (Athey and Wager 2021) to generate optimal recommendations in the form of agronomic practices that maximize potential yield gains.

# Preliminaries

```{r, message = FALSE}
load("CSISA_KVK_Public_Workspace.RData")

CSISA_KVKestim=CSISA_KVK
table(CSISA_KVKestim$SowingSchedule) 

CSISA_KVKestim$SowingSchedule=ordered(CSISA_KVKestim$SowingSchedule,levels=c("T5","T4","T3","T2","T1"))

#Create dummies for some categorical variables
library(fastDummies)
CSISA_KVKestim <- fastDummies::dummy_cols(CSISA_KVKestim, select_columns =c("VarietyClass","SoilType","CropEstablishment","Year","District"))

```


## Graphics

```{r,message = FALSE}
# Bar graphs showing percentage of farmers adopting these practices  

library(tidyverse) 
library(ggplot2)  

bar_chart=function(dat,var){   dat|>     drop_na({{var}})|>     mutate({{var}}:=factor({{var}})|>fct_infreq())|>     ggplot()+     geom_bar(aes(y={{var}}),fill="dodgerblue4")+     theme_minimal(base_size = 16) }   

sow_plot=bar_chart(CSISA_KVKestim,SowingSchedule)+labs(y="Sowing dates") 
 
sow_plot

library(ggpubr) 
library(tidyverse) 

#Sowing dates 

SowingDate_Options_Errorplot=   CSISA_KVKestim%>%   
  drop_na(SowingSchedule) %>%   
  ggerrorplot(x = "SowingSchedule", y = "GrainYield",add = "mean", error.plot = "errorbar", color="steelblue", ggtheme=theme_bw())+   
  labs(x="Sowing date options",y="Wheat yield (t/ha)")+   
  theme_minimal(base_size = 16)+
  coord_flip()  
SowingDate_Options_Errorplot 




```

## Descriptives

```{r,message = FALSE}

library(fBasics)

summ_stats <- fBasics::basicStats(CSISA_KVKestim[,c("GrainYield","VarietyClass_LDV","SoilType_Heavy","SoilType_Medium","SoilType_Low","CropEstablishment_CT","CropEstablishment_CT-line","CropEstablishment_ZT")]) 

summ_stats <- as.data.frame(t(summ_stats)) 

# Rename some of the columns for convenience 

summ_stats <- summ_stats[c("Mean", "Stdev", "Minimum", "1. Quartile", "Median",  "3. Quartile", "Maximum")] %>%   
rename("Lower quartile" = '1. Quartile', "Upper quartile"= "3. Quartile")  

summ_stats 

```

# Causal Random Forest Model

```{r,message = FALSE}
library(grf)
library(policytree)

CSISA_KVKestim_sow=subset(CSISA_KVKestim, select=c("SowingSchedule","GrainYield","VarietyClass_LDV","SoilType_Heavy","SoilType_Medium","SoilType_Low","CropEstablishment_CT","CropEstablishment_CT-line","CropEstablishment_ZT","wc2.1_30s_elev","nitrogen_0.5cm","sand_0.5cm", "soc_5.15cm","Latitude","Longitude"))

library(tidyr)
CSISA_KVKestim_sow=CSISA_KVKestim_sow %>% drop_na()


Y_cf_sowing=as.vector(CSISA_KVKestim_sow$GrainYield)
## Causal random forest -----------------

X_cf_sowing=subset(CSISA_KVKestim_sow, select=c("VarietyClass_LDV","SoilType_Heavy","SoilType_Medium","SoilType_Low","CropEstablishment_CT","CropEstablishment_CT-line","CropEstablishment_ZT","wc2.1_30s_elev","nitrogen_0.5cm","sand_0.5cm", "soc_5.15cm","Latitude","Longitude"))


W_cf_sowing <- as.factor(CSISA_KVKestim_sow$SowingSchedule)

W.multi_sowing.forest <- probability_forest(X_cf_sowing, W_cf_sowing,
  equalize.cluster.weights = FALSE,
  seed = 2
)
W.hat.multi.all_sowing <- predict(W.multi_sowing.forest, estimate.variance = TRUE)$predictions



Y.multi_sowing.forest <- regression_forest(X_cf_sowing, Y_cf_sowing,
  equalize.cluster.weights = FALSE,
  seed = 2
)

print(Y.multi_sowing.forest)
varimp.multi_sowing <- variable_importance(Y.multi_sowing.forest)
Y.hat.multi.all_sowing <- predict(Y.multi_sowing.forest, estimate.variance = TRUE)$predictions



multi_sowing.forest <- multi_arm_causal_forest(X = X_cf_sowing, Y = Y_cf_sowing, W = W_cf_sowing ,W.hat=W.hat.multi.all_sowing,Y.hat=Y.hat.multi.all_sowing,seed=2) 

varimp.multi_sowing_cf <- variable_importance(multi_sowing.forest)

multi_sowing_ate=average_treatment_effect(multi_sowing.forest, method="AIPW")
multi_sowing_ate


varimp.multi_sowing_cf <- variable_importance(multi_sowing.forest)
vars_sowing=c("VarietyClass_LDV","SoilType_Heavy","SoilType_Medium","SoilType_Low","CropEstablishment_CT","CropEstablishment_CT-line","CropEstablishment_ZT","wc2.1_30s_elev","nitrogen_0.5cm","sand_0.5cm", "soc_5.15cm","Latitude","Longitude")

## variable importance plot ----------------------------------------------------
varimpvars_sowing=as.data.frame(cbind(varimp.multi_sowing_cf,vars_sowing))
names(varimpvars_sowing)[1]="Variableimportance_sowing"
varimpvars_sowing$Variableimportance_sowing=formatC(varimpvars_sowing$Variableimportance_sowing, digits = 2, format = "f")
varimpvars_sowing$Variableimportance_sowing=as.numeric(varimpvars_sowing$Variableimportance_sowing)
varimpplotRF_sowing=ggplot(varimpvars_sowing,aes(x=reorder(vars_sowing,Variableimportance_sowing),y=Variableimportance_sowing))+
   geom_jitter(color="steelblue")+
   coord_flip()+
   labs(x="Variables",y="Variable importance")
 previous_theme <- theme_set(theme_bw(base_size = 16))
 varimpplotRF_sowing


# Policy tree --------------------------------------
DR.scores_sowing <- double_robust_scores(multi_sowing.forest)

tr_sowing <- policy_tree(X_cf_sowing, DR.scores_sowing, depth = 2) 
plot(tr_sowing)


tr_sowing3 <- hybrid_policy_tree(X_cf_sowing, DR.scores_sowing, depth = 3) 
tr_sowing3
plot(tr_sowing3)

tr_sowing4 <- hybrid_policy_tree(X_cf_sowing, DR.scores_sowing, depth = 4) 
tr_sowing4
plot(tr_sowing4)

tr_assignment_sowing=CSISA_KVKestim_sow

tr_assignment_sowing$depth2 <- predict(tr_sowing, X_cf_sowing)
table(tr_assignment_sowing$depth2)
tr_assignment_sowing$depth3 <- predict(tr_sowing3, X_cf_sowing)
table(tr_assignment_sowing$depth3)
tr_assignment_sowing$depth4 <- predict(tr_sowing4, X_cf_sowing)
table(tr_assignment_sowing$depth4)

```

# Policy learning algorithm for treatment assignment

```{r,message = FALSE}

library(rgdal)

tr_assignment_sowing$depth2_cat[tr_assignment_sowing$depth2==1]="T5_16Dec"
tr_assignment_sowing$depth2_cat[tr_assignment_sowing$depth2==2]="T4_15Dec"
tr_assignment_sowing$depth2_cat[tr_assignment_sowing$depth2==3]="T3_30Nov"
tr_assignment_sowing$depth2_cat[tr_assignment_sowing$depth2==4]="T2_20Nov"
tr_assignment_sowing$depth2_cat[tr_assignment_sowing$depth2==5]="T1_10Nov"

tr_assignment_sowingsp= SpatialPointsDataFrame(cbind(tr_assignment_sowing$Longitude,tr_assignment_sowing$Latitude),data=tr_assignment_sowing,proj4string=CRS("+proj=longlat +datum=WGS84"))

library(mapview)
mapviewOptions(fgb = FALSE)
tr_assignment_sowingspmapview=mapview(tr_assignment_sowingsp,zcol="depth2_cat",layer.name="Recommended sowing dates")
tr_assignment_sowingspmapview

```

# Distributional analysis

```{r,message = FALSE}
library(ggridges)
library(dplyr)
tau.multi_sowing.forest=predict(multi_sowing.forest, target.sample = "all",estimate.variance=TRUE)

tau.multi_sowing.forest=as.data.frame(tau.multi_sowing.forest)


tau.multi_sowing.forest_X=data.frame(CSISA_KVKestim_sow,tau.multi_sowing.forest)


# Ridges -------------------
tau.multi_sowing.forest_pred=tau.multi_sowing.forest[,1:4]

library(dplyr)
library(reshape2)
tau.multi_sowing.forest_pred=rename(tau.multi_sowing.forest_pred,"T4_15Dec - T5_16Dec"="predictions.T4_15Dec...T5_16Dec.Y.1")

tau.multi_sowing.forest_pred=rename(tau.multi_sowing.forest_pred,"T3_30Nov-T5_16Dec"="predictions.T3_30Nov...T5_16Dec.Y.1")

tau.multi_sowing.forest_pred=rename(tau.multi_sowing.forest_pred,"T2_20Nov-T5_16Dec"="predictions.T2_20Nov...T5_16Dec.Y.1")

tau.multi_sowing.forest_pred=rename(tau.multi_sowing.forest_pred,"T1_10Nov-T5_16Dec"="predictions.T1_10Nov...T5_16Dec.Y.1")


tau.multi_sowing.forest_pred_long=reshape2::melt(tau.multi_sowing.forest_pred[,1:4])

ggplot(tau.multi_sowing.forest_pred_long, aes(x=value, y=variable, fill = factor(stat(quantile)))) +
  stat_density_ridges(
    geom = "density_ridges_gradient", calc_ecdf = TRUE,
    quantiles = 4, quantile_lines = TRUE
  ) +
  scale_fill_viridis_d(name = "Quartiles")+
  theme_bw(base_size = 16)+labs(x="Wheat yield gain(t/ha)",y="Sowing date options")

```

# Transition matrix of the policy change

```{r,message = FALSE}

tr_assignment_sowing$depth2_cat[tr_assignment_sowing$depth2 == 1] <- "T5_16Dec"
tr_assignment_sowing$depth2_cat[tr_assignment_sowing$depth2 == 2] <- "T4_15Dec"
tr_assignment_sowing$depth2_cat[tr_assignment_sowing$depth2 == 3] <- "T3_30Nov"
tr_assignment_sowing$depth2_cat[tr_assignment_sowing$depth2 == 4] <- "T2_20Nov"
tr_assignment_sowing$depth2_cat[tr_assignment_sowing$depth2 == 5] <- "T1_10Nov"


library(ggalluvial)
library(data.table)
tr_assignment_sowingDT = data.table(tr_assignment_sowing)
TransitionMatrix_sowing <- tr_assignment_sowingDT[, (sum <- .N), by = c("SowingSchedule", "depth2_cat")]
library(dplyr)
TransitionMatrix_sowing <- rename(TransitionMatrix_sowing, Freq = V1)

library(scales)
transitionmatrixplot_sowing <- ggplot(
    data = TransitionMatrix_sowing,
    aes(axis1 = SowingSchedule, axis2 = depth2_cat, y = Freq)
) +
    geom_alluvium(aes(fill = depth2_cat)) +
    geom_stratum() +
    # geom_text(stat="stratum", aes(label=after_stat(stratum),nudge_y =5))+
    geom_text(stat = "stratum", aes(label = paste(after_stat(stratum), percent(after_stat(prop))))) +
    scale_x_discrete(
        limits = c("SowingSchedule", "depth2_cat"),
        expand = c(0.15, 0.05)
    ) +
    scale_fill_viridis_d() +
    theme_void(base_size = 20) +
    theme(legend.position = "none")

transitionmatrixplot_sowing


```
